Causal Forest Approach for Site-specific Input Management via On-farm Precision Experimentation

### 2022-02-14 Writing/appendix.rmd
New results and added training accuracy of the ML methods to supplementary

```{r ct-ex, fig.width=5, fig.height=5, fig.align="center", fig.cap = "An illustrative example of causal tree"}
fancyRpartPlot(opfit, sub = "")
```

@@

# Appendix B: Understanding yield response functions better

@@

\begin{alinged}
f(N, \mathbf{c}) = ymax(\mathbf{c})(1-exp(\alpha(\mathbf{c}) + \beta(\mathbf{c}) \cdot N)) + \varepsilon. 
\end{alinged}

@@

# Appendix C: Parameter generation

@@

```{r variogram-parameters, tab.cap = "The Parameters for the Variogram Models"}
variogram_tb
```

```{r fig-parameters, fig.width=7, fig.height=9, fig.cap = "An Illustrative Example of Spatial Distributions of Field Characteristics", cache = TRUE}

@@

## Splitting the base parameters {#split-pars}

@@

>>>>

```{r ct-ex, fig.width=5, fig.height=5, fig.align="center", fig.cap = "An illustrative example of a causal tree"}
fancyRpartPlot(opfit, sub = "")
```

@@

# Appendix B: Understanding Yield Response Functions Better

@@

\begin{equation}
f(N, \mathbf{c}) = ymax(\mathbf{c})(1-exp(\alpha(\mathbf{c}) + \beta(\mathbf{c}) \cdot N)) + \varepsilon. 
\end{equation}

@@

# Appendix C: Parameter Generation

@@

```{r variogram-parameters, tab.cap = "The parameters for the Variogram models"}
variogram_tb
```

```{r fig-parameters, fig.width=7, fig.height=9, fig.cap = "An Illustrative wxample of spatial distributions of field characteristics", cache = TRUE}

@@

## Splitting the Base Parameters {#split-pars}

@@

# Appendix C: Training Accuracy of the ML methods

During the simulations, we evaluated the training accuracy using the same ML models we presented in the Result and Discussion section. 
Table \@ref(tab:table-optN-train) shows the mean RMSE of the EONR estimation and $\hat{\pi}_{def}$ over the one thousand simulations using the training datasets as testing datasets. Overall, the performances of each ML method did not change between training and testing datasets. Regarding CNN, the identified site-specific yield response functions were perfectly linear as we confirmed in the testing results.

```{r table-optN-train, tab.cap = "Mean RMSE of EONR (kg/ha) estimation and profit-deficit ($/ha) by ML methods and modeling scenarios on training datasets"}
report_table_optN_train
```

Table \@ref(tab:table-y-train) shows the mean RMSE of yield prediction over the one thousand simulations using the training datasets as testing datasets. Interestingly, the performances of the prediction-oriented ML methods in predicting yields were relatively improved compared with the test results. Despite the improvement of prediction-oriented ML methods in yield prediction accuracy, it did not result in the more accurate EONR estimates. This supports our argument that the model’s yield prediction accuracy has little to do with its EONR prediction accuracy.

```{r table-y-train, tab.cap = "Mean RMSE of yield prediction (kg/ha) by ML methods and modeling scenarios on training datasets"}
report_table_y_train
```

### 2022-02-14 Writing/manuscript_cea.rmd
minor changes
New results and added training accuracy of the ML methods to supplementary

Monte Carlo simulations were conducted for nitrogen experiments for corn production to estimate site-specific EONRs. The simulations' first step generated one thousand synthetic fields that "resembled" an actual field. Field characteristics and yield were simulated assuming the Mitscherlich-Baule (MB) yield response function. We used the RF, BRF, and CNN ML methods to estimate yield response functions, and we used the CF method to estimate treatment effects. For each method, we estimated site-specific EONRs using the training data set, and checked the statistical performance of the predicted site-specific EONR using the validation data set. Detailed descriptions of these steps are provided below. The *R* and *python* codes that implement the simulation analyses are available at @Kakimoto2022codes.

@@

A cell-level data set for the whole field, $\{(y_{i,j}, N^{opt}_{i,j}, N_{i,j}, \mathbf{c}_{i,j}): i \in \{1,2,\dots,1536\}, j \in \{1,2,\dots,36\}\}$ was generated by plugging in the parameter values, the error term, and the assigned nitrogen rates into \@ref(eq:eqn6). The cell-level data were then aggregated up to the subplot level by taking the mean of the cell-level values inside each subplot, which were the unit of observation. The subplot data were therefore

@@

In subsequent modeling scenarios, we made the simulated experimental data more realistic by varying the degree to which the researcher understood what the needed covariates' values were those in the data set. In the second scenario, denoted "abytt," we introduced additional variables $\theta^1$ and $\theta^2$, and the researcher's data set contained values of $\alpha_{i}$, $\beta_{i}$, $ymax_{i}$, $\theta^1_{i}$ and $\theta^2_{i}$ for all subplots $i=1,\dots,1536$. $\theta^1$ and $\theta^2$ had no effect on yield, but they were spatially correlated with $\beta_{i}$. This scenario was meant to reflect the lack of scientific consensus about which variables explain heterogeneous yield response to N, making it likely that the researcher would include irrelevant variables like $\theta^1$ and $\theta^2$ in real-world data analysis.

In the third scenario, denoted "aabbyy," the researcher's data set did not contain values of $\alpha_{i}$, $\beta_{i}$, or $ymax_{i}$. Rather, it contained subplot-level values of permutations of $\alpha$, $\beta$, and $ymax$. Each of the three parameters were split into two parts in a spatially correlated manner, and six new covariates were created (see Appendix \@ref(split-pars) for the details about this parameter-splitting process). These were $\alpha^{1}_{i}$, $\alpha^{2}_{i}$, $\beta^{1}_{i}$, $\beta^{2}_{i}$, $ymax^{1}_{i}$, and $ymax^{2}_{i}$. The two split variables were spatially correlated to reflect that many of the soil/field characteristics used to analyze OFPE data are spatially correlated. The values of $\alpha_{i}$, $\beta_{i}$ and $ymax_{i}$ did not indicate specific soil or field characteristics, but were functions of such. In reality, scientists cannot directly observe variables that can accurately predict yield plateau level, soil N content and N uptake efficiency. Rather, they use multiple observed soil/field characteristics to explain such phenological phenomena. For example, instead of using $ymax_{i}$ directly, some topographical land features and soil properties such as elevation and soil sand content could be used as yield-limiting factors [@jiang2004effect; @kravchenko2000correlation]. In reality, scientists may include more than three variables as covariates. This scenario should reduce the accuracy of EONR modeling for all the models compared to the ideal case.

@@

The RF, BRF, and CNN methods all followed the same conceptual steps to estimate site-specific EONR: 1) estimate the yield production functions and 2) calculate site-specific EONRs based on that estimated function. For the given modeling scenario, RF and BRF predict yield by using all the available covariates as explanatory variables in the estimation process. The `grf` package (version 1.2.0 [@tibshirani2018package]) in R was used for RF and BRF modeling. RF ann BRF from the `grf` package have several tunable hyperparameters (e.g., the minimum node size in each tree, the number of covariates used for node splitting, and parameters involving an honest tree-building process), and the optimal values for these parameters were selected by cross-validation. Individual trees in a forest were built with randomly subsampled observations, leading to a different prediction in every forest even when the same data set was used. To reduce the variance of predictions for prediction accuracy, we set the number of trees (*num.trees*) at a high value, $4000$.

@@

For evaluation of yield and EONR predictions, a data set from a different simulation round was used as the test data set. For example, for simulation number 1, the training data set from simulation number 2 was used as the test data set. This process works because the training data sets were generated by the same data generating process, yet have different underlying field parameters ($\alpha$, $\beta$, $ymax$). Further, all training data sets were generated independently, so for any particular simulation round, the data set from any other round could serve as a valid test data set. This practice made the creation of one thousand additional simulations fields unnecessary.

@@

Table \@ref(tab:table-optN-subplot-wide) shows the mean RMSE of the EONR estimation and $\hat{\pi}_{def}$ over the one thousand simulations; it makes immediately clear that CF-base considerably outperformed the other ML methods, and that the profits based on CF-base's EONR estimates were the closest to the true maximum profits. BRF performed considerably better predicting EONR than did RF.

@@

```{r table-optN-subplot-wide, tab.cap = "Mean RMSE of EONR (kg/ha) estimation and profit-deficit ($/ha) by ML methods and modeling scenarios"}
report_table_optN
```

@@

Table \@ref(tab:table-y-subplot-wide) shows the mean RMSE of yield prediction over the one thousand simulations. Since CF-base does not predict yield, the table only reports the results from RF, BRF and CNN. Compared to the result of EONR estimation, RF performed as well as BRF in the simple modeling scenarios of "aby" and "abytt." This fact along with the poor performance of RF in predicting EONR show that strong predictive power of yield levels does not imply accurate prediction of EONR. Further, RF shows vulnerability to the inclusion of many correlated variables, as is displayed for EONR prediction as well. While CNN performed worse than RF and BRF in "aby" and "abytt," it performed better than RF in more complicated modeling scenarios, showing robustness, at least in yield prediction, to inclusion of many correlated variables. Of course, while on average in the "aabbyy" and "aabbyytt" scenarios CNN predicted yields better than RF did, CNN's EONR predictions were worse than RF's, which further backs our claim that a good yield prediction capability does not imply good EONR prediction capability.

@@

```{r table-y-subplot-wide, tab.cap = "Mean RMSE of yield prediction (kg/ha) by ML methods and modeling scenarios"}
report_table_y
```

>>>>

Monte Carlo simulations were conducted for nitrogen experiments for corn production to estimate site-specific EONRs. The simulations' first step generated one thousand synthetic fields that "resembled" an actual field. Field characteristics and yield were simulated assuming the Mitscherlich-Baule (MB) yield response function. We used the RF, BRF, and CNN ML methods to estimate yield response functions, and we used the CF method to estimate treatment effects. For each method, we estimated site-specific EONRs using the training dataset, and checked the statistical performance of the predicted site-specific EONR using the validation dataset. Detailed descriptions of these steps are provided below. The *R* and *python* codes that implement the simulation analyses are publicly accessible as a Mendeley dataset [@Kakimoto2022codes].

@@

A cell-level dataset for the whole field, $\{(y_{i,j}, N^{opt}_{i,j}, N_{i,j}, \mathbf{c}_{i,j}): i \in \{1,2,\dots,1536\}, j \in \{1,2,\dots,36\}\}$ was generated by plugging in the parameter values, the error term, and the assigned nitrogen rates into \@ref(eq:eqn6). The cell-level data were then aggregated up to the subplot level by taking the mean of the cell-level values inside each subplot, which were the unit of observation. The subplot data were therefore

@@

In subsequent modeling scenarios, we made the simulated experimental data more realistic by varying the degree to which the researcher understood what the needed covariates' values were those in the dataset. In the second scenario, denoted "abytt," we introduced additional variables $\theta^1$ and $\theta^2$, and the researcher's dataset contained values of $\alpha_{i}$, $\beta_{i}$, $ymax_{i}$, $\theta^1_{i}$ and $\theta^2_{i}$ for all subplots $i=1,\dots,1536$. $\theta^1$ and $\theta^2$ had no effect on yield, but they were spatially correlated with $\beta_{i}$. This scenario was meant to reflect the lack of scientific consensus about which variables explain heterogeneous yield response to N, making it likely that the researcher would include irrelevant variables like $\theta^1$ and $\theta^2$ in real-world data analysis.

In the third scenario, denoted "aabbyy," the researcher's dataset did not contain values of $\alpha_{i}$, $\beta_{i}$, or $ymax_{i}$. Rather, it contained subplot-level values of permutations of $\alpha$, $\beta$, and $ymax$. Each of the three parameters were split into two parts in a spatially correlated manner, and six new covariates were created (see Appendix \@ref(split-pars) for the details about this parameter-splitting process). These were $\alpha^{1}_{i}$, $\alpha^{2}_{i}$, $\beta^{1}_{i}$, $\beta^{2}_{i}$, $ymax^{1}_{i}$, and $ymax^{2}_{i}$. The two split variables were spatially correlated to reflect that many of the soil/field characteristics used to analyze OFPE data are spatially correlated. The values of $\alpha_{i}$, $\beta_{i}$ and $ymax_{i}$ did not indicate specific soil or field characteristics, but were functions of such. In reality, scientists cannot directly observe variables that can accurately predict yield plateau level, soil N content and N uptake efficiency. Rather, they use multiple observed soil/field characteristics to explain such phenological phenomena. For example, instead of using $ymax_{i}$ directly, some topographical land features and soil properties such as elevation and soil sand content could be used as yield-limiting factors [@jiang2004effect; @kravchenko2000correlation]. In reality, scientists may include more than three variables as covariates. This scenario should reduce the accuracy of EONR modeling for all the models compared to the ideal case.

@@

The RF, BRF, and CNN methods all followed the same conceptual steps to estimate site-specific EONR: 1) estimate the yield production functions and 2) calculate site-specific EONRs based on that estimated function. For the given modeling scenario, RF and BRF predict yield by using all the available covariates as explanatory variables in the estimation process. The `grf` package (version 1.2.0 [@tibshirani2018package]) in R was used for RF and BRF modeling. RF ann BRF from the `grf` package have several tunable hyperparameters (e.g., the minimum node size in each tree, the number of covariates used for node splitting, and parameters involving an honest tree-building process), and the optimal values for these parameters were selected by cross-validation. Individual trees in a forest were built with randomly subsampled observations, leading to a different prediction in every forest even when the same dataset was used. To reduce the variance of predictions for prediction accuracy, we set the number of trees (*num.trees*) at a high value, $2000$.

@@

For evaluation of yield and EONR predictions, a dataset from a different simulation round was used as the test dataset. For example, for simulation number 1, the training dataset from simulation number 2 was used as the test dataset. This process works because the training datasets were generated by the same data generating process, yet have different underlying field parameters ($\alpha$, $\beta$, $ymax$). Further, all training datasets were generated independently, so for any particular simulation round, the dataset from any other round could serve as a valid test dataset. This practice made the creation of one thousand additional simulations fields unnecessary.

@@

Table \@ref(tab:table-optN) shows the mean RMSE of the EONR estimation and $\hat{\pi}_{def}$ over the one thousand simulations; it makes immediately clear that CF-base considerably outperformed the other ML methods, and that the profits based on CF-base's EONR estimates were the closest to the true maximum profits. BRF performed considerably better predicting EONR than did RF.

@@

```{r table-optN, tab.cap = "Mean RMSE of EONR (kg/ha) estimation and profit-deficit ($/ha) by ML methods and modeling scenarios"}
report_table_optN
```

@@

Table \@ref(tab:table-y) shows the mean RMSE of yield prediction over the one thousand simulations. Since CF-base does not predict yield, the table only reports the results from RF, BRF and CNN. Compared to the result of EONR estimation, RF performed as well as BRF in the simple modeling scenarios of "aby" and "abytt." This fact along with the poor performance of RF in predicting EONR show that strong predictive power of yield levels does not imply accurate prediction of EONR. Further, RF shows vulnerability to the inclusion of many correlated variables, as is displayed for EONR prediction as well. While CNN performed worse than RF and BRF in "aby" and "abytt," it performed better than RF in more complicated modeling scenarios, showing robustness, at least in yield prediction, to inclusion of many correlated variables. Of course, while on average in the "aabbyy" and "aabbyytt" scenarios CNN predicted yields better than RF did, CNN's EONR predictions were worse than RF's, which further backs our claim that a good yield prediction capability does not imply good EONR prediction capability.

@@

```{r table-y, tab.cap = "Mean RMSE of yield prediction (kg/ha) by ML methods and modeling scenarios"}
report_table_y
```
